\vspace{-0.5cm} XSGRB - The X-shooter GRB afterglow sample

### 2018-08-28 document/referee_response.txt

This report presents a review of the article "The X-shooter GRB 
afterglow legacy sample (XS-GRB)" by Selsing and collaborators. The 
authors present and make available to the community a complete sample 
of GRB afterglows observed with the X-Shooter spectrograph on the 
VLT. This dataset is remarkable in several respect (>100 spectra 
available taken over 8 years) and some of these have not be published 
elsewhere. However, there are a number of important issues regarding 
both the scientific measurements and their presentation which are 
detailed below. These concerns need to be addressed before the paper 
can be considered for publication. More details on these points 
together with other comments are provided below.

Major Concerns:

# - One aspect of these data that needs more emphasis is the 
# intervening absorbers. The data presented in this paper present a 
# remarkable number of metal and HI intervening absorbers as can be 
# judged from reading the appendix. These deserve much more attention, 
# including a table listing their redshifts, column density 
# measurements when available, metal lines detected, etc. A figure 
# with their redshift distribution would also be very welcome. In 
# addition, it would be interesting to compare the properties of this 
# population with the absorbers towards background quasars.

While we agree that the subject of intervening absorbers is very interesting and
important, the main focus of this paper is to present the technicalities of the
sample and the data reduction. In a series of fourth-coming papers, we will
investigate sub-themes independently. On the subject of intervening absorbers,
already published is Christensen et al., A&A, 608A, 84C, 2017,
(arXiv:1709.01084) and we also have de Ugarte Postigo et al. (in prep).

# - The manuscript states that the sample is homogeneous which is a bit 
# misleading. The spectrograph used to make the follow-up is the same 
# for all afterglows, but the SNR (due to a combination of brightness 
# of the source and time of observations after the burst) and the 
# rest-frame wavelength coverage vary enormously. While remarkable in 
# many respects, it is not appropriate to qualify this sample as 
# homogeneous.

We completely agree and have rephrased the manuscript to reflect this change.

# - There are a great many details provided about the data reduction, some of 
# which the reader fails to relate to the final scientific goals. For 
# example, a correction is applied to correct offsets in the 
# wavelength calibration, which presumably mostly impact the redshift 
# determination. How does the improvement translate into the redshift 
# determination accuracy? Similarly, for the telluric correction. What 
# is the impact of this on the final science? Conversely, a 
# description of the flux calibration error estimate would have been 
# welcome.

# - The comparison of N(HI) distributions between intervening quasar 
# absorbers and intrinsic GRB absorbers is not direct. The measurement 
# of the column density of a system in the case of a quasar absorber 
# integrate across the whole system. Conversely, in the case of GRBs, 
# the burst event takes place at a random position in the host galaxy 
# and in fact, on average, the column density will measure over only 
# half of the system. This geometric effect should be taken into 
# account.

# - Another concern is to do with using absorption features to measure 
# the redshift. Indeed, one cannot exclude that the absorption is not 
# at the position of host galaxy but lying in front of it. For this 
# reason, the redshifts provided should be seen as upper limits.

Other Concerns:

- the links to the actual data seem to be inactive. It is important to 
be able to review the material which is made available to the 
community.

- caption of Fig. 3 is cut-out

- Fig. 5: please label the y-axis

- Fig. 10: caption is cut-out again 
GRB150403A: could the continuum be lower? 
GRB150915A: is this fit significant?

- appendix A: refer to the spectral PSF in the title?

- appendix B: science-verified, Paranalization: 
perhaps avoid this ESO jargon?

typos: 
- pt. 4 p. 3: Paranl 
- p. 3: exampl 
- p. 3: sorrounding 
- p. 9: curios 
- there are numerous typos in the appendix which I will not review 
here. Please make sure to double check spelling there.

I would be happy to review an updated version of the manuscript where 
the above comments are taken into account.

### 2018-08-28 document/XSGRB_sample_arxiv.tex
Arxiv submission

@@

\usepackage{url}
\usepackage{color,hyperref}
\definecolor{linkcolor}{rgb}{0,0.3,0.7}
\hypersetup{colorlinks=true,
	linkcolor=linkcolor, 
	citecolor=linkcolor,
	filecolor=linkcolor, 
	urlcolor=linkcolor}

\usepackage{natbib}
%\usepackage{caption}

@@

	\item No bright, nearby stars within $ 1.8 + 0.4 \times \exp[(20 - R)/2.05$]
arcsec, where $R$ is the USNO magnitude of the star.

@@

\begin{figure*}
	\centerline{\includegraphics[width=\linewidth]{figures/GRB121024A.pdf}}
\caption{Telluric corrected, normalized spectrum of GRB~121024A at $z = 2.300$
	that illustrates the typical data quality. The continuum estimate is shown in
	dashed red and the error spectrum in solid blue. The acquisition magnitude is
	$R$ = 20, meaning it is in the brighter end of the sample presented here, but
	not the brightest. The spectrum is rich in absorption and emission lines, including
	absorption from molecular $\mathrm{H_2}$. The absorption trough visible at
	$\sim$ 4000 $\AA$ is due to \lya~in the host. We have marked the most prominent
	lines seen in GRB afterglows from \citet{Christensen2011a}. The regions of most
	severe telluric absorption are highlighted by grey-shading the background.
	Additionally, three intervening systems are seen in this sightline. This
	spectrum has been analyzed in detail in \citet{Friis2015}.}
\label{fig:spectrum}
\end{figure*}

@@

All the spectra are made available as a single
\href{https://sid.erda.dk/cgi-sid/ls.py?share_id=DBuNORk1lI}{ZIP file}, through
\url{http://grbspec.iaa.es} \citep{DeUgartePostigo2014f}, and additionally through the ESO archive in the
form of \href{http://archive.eso.org/wdb/wdb/adp/phase3_main/form}{phase 3
	material}. This release includes both prompt afterglow observations as well as
late time observations of the associated hosts, and represents \textit{all}
afterglow spectra of GRBs carried out by the X-shooter spectrograph since the
commissioning of the instrument, \startdate, and until the end of the last
period of the program 098.A-0055, \termdate~and thus constitutes eight years of GRB
afterglow observations with X-shooter. An overview of all the spectra and their
observational setups is given in Table \ref{tab:sample_overview}. For each
burst, each individual observation is provided in a separate reduction, and in
cases where observations have been repeated for an increased signal-to-noise or
to follow the temporal evolution, a combined spectrum is also provided. No
attempt has been made to join the spectroscopic arms, so for each observation,
three spectra are provided in separate files.

@@

\bibliographystyle{aa}
\bibliography{XSGRB_sample,GCN}

>>>>

%\usepackage{hyperref}

@@

\usepackage{color}
\definecolor{linkcolor}{rgb}{0,0.3,0.7}
\hypersetup{colorlinks=true,
	linkcolor=linkcolor, 
	citecolor=linkcolor,
	filecolor=linkcolor, 
	urlcolor=linkcolor}

\usepackage{url}
%\usepackage{caption}

@@

	\item No bright, nearby stars within $ 1.8 + 0.4 \times \exp[(20 - R)/2.05$]
arcsec, where $R$ is the USNO magnitude of the star).

@@

\begin{figure*}
	\centerline{\includegraphics[width=16cm]{figures/GRB121024A.pdf}}
\caption{Telluric corrected, normalized spectrum of GRB~121024A at $z = 2.300$
	that illustrates the typical data quality. The continuum estimate is shown in
	dashed red and the error spectrum in solid blue. The acquisition magnitude is
	$R$ = 20, meaning it is in the brighter end of the sample presented here, but
	not the brightest. The spectrum is rich in absorption and emission lines, including
	absorption from molecular $\mathrm{H_2}$. The absorption trough visible at
	$\sim$ 4000 $\AA$ is due to \lya~in the host. We have marked the most prominent
	lines seen in GRB afterglows from \citet{Christensen2011a}. The regions of most
	severe telluric absorption are highlighted by grey-shading the background.
	Additionally, three intervening systems are seen in this sightline. This
	spectrum has been analyzed in detail in \citet{Friis2015}.}
\label{fig:spectrum}
\end{figure*}

@@

All the spectra are made available as a single
\href{https://sid.erda.dk/cgi-sid/ls.py?share_id=DBuNORk1lI}{ZIP file}, through
\url{http://grbspec.iaa.es} \citep{deUgartePostigo2014f}, and additionally through the ESO archive in the
form of \href{http://archive.eso.org/wdb/wdb/adp/phase3_main/form}{phase 3
	material}. This release includes both prompt afterglow observations as well as
late time observations of the associated hosts, and represents \textit{all}
afterglow spectra of GRBs carried out by the X-shooter spectrograph since the
commissioning of the instrument, \startdate, and until the end of the last
period of the program 098.A-0055, \termdate~and thus constitutes eight years of GRB
afterglow observations with X-shooter. An overview of all the spectra and their
observational setups is given in Table \ref{tab:sample_overview}. For each
burst, each individual observation is provided in a separate reduction, and in
cases where observations have been repeated for an increased signal-to-noise or
to follow the temporal evolution, a combined spectrum is also provided. No
attempt has been made to join the spectroscopic arms, so for each observation,
three spectra are provided in separate files.

@@

\bibliographystyle{aa_arxiv}
\bibliography{XSGRB_sample,GCN}
